---
title: "R Notebook"
output: html_notebook
---

Read in data

```{r}
foo <- data.table::fread('data/fitness_freshness.csv',data.table=F)

foo$Date <- as.Date(foo$Date,format="%m/%d/%y")
foo$diffFitness <- c(NA,diff(foo$Fitness))
foo$diffFatigue <- c(NA,diff(foo$Fatigue))
```


plots

```{r}

g <- ggplot2::ggplot(foo,ggplot2::aes(Date,Fitness))
g <- g+ggplot2::geom_smooth()
g <- g+ggplot2::geom_point()
g

g <- ggplot2::ggplot(foo,ggplot2::aes((Impulse),(diffFitness)))
g <- g+ggplot2::geom_point()
g

g <- ggplot2::ggplot(foo,ggplot2::aes(Date,diffFitness))
#g <- g+ggplot2::geom_smooth()
g <- g+ggplot2::geom_point()
g

g <- ggplot2::ggplot(foo,ggplot2::aes(Date,log10(Impulse)))
#g <- g+ggplot2::geom_smooth()
g <- g+ggplot2::geom_point()
g
g <- ggplot2::ggplot(foo,ggplot2::aes((Impulse),(diffFatigue)))
g <- g+ggplot2::geom_point()
g

g <- ggplot2::ggplot(foo,ggplot2::aes(Date,Fatigue))
g <- g+ggplot2::geom_smooth()
g <- g+ggplot2::geom_point()
g

g <- ggplot2::ggplot(foo,ggplot2::aes(Date,Form))
g <- g+ggplot2::geom_smooth()
g <- g+ggplot2::geom_point()
g

foo2 <- dplyr::filter(foo,Impulse==0)

g <- ggplot2::ggplot(foo2,ggplot2::aes(Fitness,diffFitness))
g <- g+ ggplot2::geom_point()
g

g <- ggplot2::ggplot(foo2,ggplot2::aes(Fitness,diffFatigue))
g <- g+ ggplot2::geom_point()
g


```



```{r}



new_fitness_freshness <- function(x,y,z){
  transitionKernel <- rbind(c(-.21,.979,0,0.02),c(1.51,-0.19,1,.139))
  return(transitionKernel%*%(c(1,x,y,z)))
}

run_three_months <- function(xinit,yinit,impulse){
  res <- list()
  res$fitness <- c()
  res$fatigue <- c()
  x <- new_fitness_freshness(xinit,yinit,0)
  for (i in 1:26){
    for(j in 1:14){
      x <- new_fitness_freshness(x[1],x[2],impulse[j])
      res$fitness <- c(res$fitness,x[1])
      res$fatigue <- c(res$fatigue,x[2])
    }
  }
  return(res)
}

res_simple <- run_three_months(63,53,rep(62.90541,14))

```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

