---
title: "Load gpx"
output: html_notebook
---

Process gpx file.

```{r}
read_gpx_file <- function(filename){
  tmt2 <- xml2::read_xml(filename)
  tmt2b<-xml2::as_list(tmt2)
  #lat/lon/elev/time
  lat <- sapply(tmt2b$gpx$trk$trkseg,attr,'lat')
  lon <- sapply(tmt2b$gpx$trk$trkseg,attr,'lon')
  elev <- sapply(tmt2b$gpx$trk$trkseg,function(x) as.numeric(x$ele[[1]]))
  time <- sapply(tmt2b$gpx$trk$trkseg,function(x) as.character(x$time[[1]]))
  
  #hr/cad/temp/power
  hr <- sapply(tmt2b$gpx$trk$trkseg,
               function(x) as.numeric(x$extensions$TrackPointExtension$hr[[1]][[1]]))
  cad <- sapply(tmt2b$gpx$trk$trkseg,
                function(x) as.numeric(x$extensions$TrackPointExtension$cad[[1]][[1]]))
  temp <- sapply(tmt2b$gpx$trk$trkseg,
                 function(x) as.numeric(x$extensions$TrackPointExtension$atemp[[1]][[1]]))
  power <- sapply(tmt2b$gpx$trk$trkseg,
                  function(x) as.numeric(x$extensions$power[[1]]))

  #missingness
  fxn1 <- function(x){
    vec1 <- sapply(x,length)==0
    n1 <- sum(vec1)
    if(n1>0){
      x[vec1] <- NA
      x <- sapply(x, function(x) x)
      return(x)
    } else {
      return(x)
    }
  }
  hr <- fxn1(hr)
  cad <- fxn1(cad)
  temp <- fxn1(temp)
  power <- fxn1(power)
  df <- data.frame(time=time,
                   latitude=as.numeric(lat),
                   longitude=as.numeric(lon),
                   elevation=elev,
                   hr=hr,
                   cad=cad,
                   temp=temp,
                   power=power,
                   stringsAsFactors = F)
  
  cat('computing smoothed distances\n')
  secs <- nrow(df)
  geo_dist <- geosphere::distVincentyEllipsoid(cbind(df$longitude[-secs],
                                                     df$latitude[-secs]),
                                               cbind(df$longitude[-1],
                                                     df$latitude[-1]))
  geo_dist <- c(0,smooth::sma(geo_dist)$fitted)
  elev_dist <- df$elevation[-1]-df$elevation[-secs]
  elev_dist <- c(0,smooth::sma(elev_dist)$fitted)
  df$dist <- sqrt(geo_dist^2+elev_dist^2)
  df$geo_dist <- geo_dist
  df$ascent <- rep(0,secs)
  df$descent <- rep(0,secs)
  df$ascent[elev_dist>0] <- elev_dist[elev_dist>0]
  df$descent[elev_dist<0] <- elev_dist[elev_dist<0]
  df$time <- sapply(df$time,lubridate::as_datetime,tz='America/Los_Angeles')
  df$elapsedtime <- c(0,df$time[-1]-df$time[-secs])
  df$speed <- c(0,(df$elapsedtime[-1]/df$dist[-1])*26.8224)
  df$speed <- smooth::sma(df$speed)$fitted
  return(df)
}

```

play around with some data
```{r message=FALSE}
fun_run <- read_gpx_file('data/gpx_tracks/Fun_run.gpx')
fun_run_imp <- bcv::impute.svd(fun_run)$x
colnames(fun_run_imp) <- colnames(fun_run)

fun_run_imp <- fun_run_imp[fun_run_imp[,'cad']>0,]
umap_res<-umap::umap((fun_run_imp[,c('elevation',
                                     'hr',
                                     'cad',
                                     'temp',
                                     'power',
                                     'speed')]))

umapDf <- data.frame(dim1=umap_res$layout[,1],
                     dim2=umap_res$layout[,2],
                     stringsAsFactors = F)
umapDf <- cbind(umapDf,fun_run[fun_run$cad>0,])
umapDf <- data.frame(umapDf,stringsAsFactors = F)

ab<-kmeans(umapDf[,c('dim1','dim2')],10,nstart=1000)
umapDf$cluster <- as.factor(ab$cluster)
g <- ggplot2::ggplot(umapDf,
                     ggplot2::aes(dim1,
                                  dim2,
                                  color=cluster,
                                  alpha=elevation))

g <- g+ggplot2::geom_point()
g

h <- ggplot2::ggplot(umapDf,
                     ggplot2::aes(longitude,
                                  latitude,
                                  color=cluster))

h <- h+ggplot2::geom_jitter()
h
```

ideas to follow up on

1. fix speed given pause in data recording
2. fix time formatting - done
3. run umap on data to see if there are interesting patterns
4. extract features (mean, variance, sd, skew, kurtosis...???)
5. figure out efficient way to download all of gpx files going back to deception pass 50k, and set up a function to process all of them to make unified data-set
6. 
