---
title: "Ultra training notebook"
output: html_notebook
---

Load data and convert date

```{r}
foo <- data.table::fread('data/hundo_training.csv')
foo$Date <- as.Date(foo$Date,tryFormats=c("%m/%d/%y"))
```

Create beats per mile and beats per grade adjusted mile stats

```{r}
foo$bpgam <- foo$GAP*foo$HR
foo$bpmile <- foo$Pace*foo$HR
```

Create a numeric Date variable starting at the commencement of serious training (April 9, 2019), as well as a rest days variabe.

```{r}
foo$DateNew <- as.numeric(foo$Date)-min(as.numeric(foo$Date))
foo$restDays <- c(8,diff(foo$DateNew))
```

Fit model without effort or DateNew to see progression.

```{r}
model1 <- lm(bpmile ~ Vert +
               log10(Distance) +
               Cadence+log10(Elevation) + 
               Vert*log10(Distance) + 
               Cadence*Vert + 
               Track + 
               restDays +
               MaxHR,foo)
foo$resid <- model1$residuals
g <- ggplot2::ggplot(foo,ggplot2::aes(x=Date,y=resid))

g <- g + ggplot2::geom_smooth()

g <- g + ggplot2::geom_point()

g
```

Look at beats per grade adjusted mile

```{r}
summary(lm(bpgam ~ DateNew,foo))
g <- ggplot2::ggplot(foo,ggplot2::aes(x=Date,
                                      y=(bpgam)))

g <- g + ggplot2::geom_smooth()

g <- g + ggplot2::geom_point()

g
```


Fit model without effort to see effects of each variable.
```{r}
model2 <- lm(bpmile ~ Vert + 
               log10(Distance) + 
               Cadence + 
               log10(Elevation) + 
               Vert*log10(Distance) + 
               Cadence*Vert + 
               Track + 
               DateNew +
               restDays + 
               MaxHR,foo)
summary(model2)
foo$residFull <- scale(model2$residuals)
```


Fit model with effort to see full effects
```{r}
model3 <- lm(bpmile ~ Vert + 
               log10(Distance) + 
               Cadence +
               log10(Elevation) + 
               Vert*log10(Distance) + 
               Cadence*Vert + 
               Track + 
               log10(Effort) + 
               DateNew +
               restDays + 
               MaxHR,foo)
summary(model3)
foo$residEffort <- scale(model3$residuals)
foo$pEffort <- pchisq(foo$residEffort^2,1,lower.tail=F)
```

Look at the full matrix

```{r}
dplyr::arrange(foo,dplyr::desc(Date))
```


compute some cumulative efforts

```{r}
fxn2 <- function(day,df){
  searchDays <- day - c(1:21)
  ab <- dplyr::filter(df,Date%in%searchDays)
  if(nrow(ab)>0){
    return(sum(ab$Distance))
  }else{
    return(0)
  }
}

foo$cumulativeDistance <- sapply(foo$Date,fxn2,foo)
modelx <- lm(bpmile ~ Vert + 
               log10(Distance) + 
               Cadence +
               log10(Elevation) + 
               Vert*log10(Distance) + 
               Cadence*Vert + 
               Track + 
               DateNew +
               restDays + 
               MaxHR + 
               log10(cumulativeDistance+1),foo)
summary(modelx)
```


Define a pace prediction function
```{r}
pacePredictor <- function(modFit,
                                DateNew=268,
                                HR=170,
                                Vert=100,
                                Distance=26.2,
                                Cadence=170,
                                Elevation=100,
                                Track=FALSE,
                                restDays=1,
                                Effort = NULL){
  if(!is.null(Effort)){
    predDf=data.frame(DateNew=DateNew,
                      HR=HR,
                      Vert=Vert,
                      Distance=Distance,
                      Cadence=Cadence,
                      Elevation=Elevation,
                      Track=Track,
                      restDays=restDays,
                      Effort=Effort,stringsAsFactors=F)
  }else{
    predDf=data.frame(DateNew=DateNew,
                      HR=HR,
                      Vert=Vert,
                      Distance=Distance,
                      Cadence=Cadence,
                      Elevation=Elevation,
                      Track=Track,
                      restDays=restDays,stringsAsFactors=F)
  }
  res <- predict(modFit,
                 predDf,
                 interval='prediction',
                 level = .68)
  names(res) <- c('Prediction','Lower Bound','Upper Bound')
  return(res)
}
```

plot run types

```{r}
set.seed(1)
dM <- dplyr::select(foo,
                    DateNew,
                    Vert,
                    Distance,
                    Cadence,
                    Elevation,
                    Effort,
                    Track,
                    Temp,
                    HR,
                    MaxHR,
                    StartTime,
                    ElapsedTime,
                    restDays)

dM$Distance <- log10(dM$Distance)
dM$Elevation <- log10(dM$Elevation)
dM$Effort <- log10(dM$Effort)
res2 <- umap::umap(scale(dM))




umapDf <- data.frame(dim1=res2$layout[,1],
                     dim2=res2$layout[,2],
                     stringsAsFactors = F)

umapDf <- cbind(umapDf,foo)

ab<-kmeans(umapDf[,c('dim1','dim2')],4,nstart=1000)
umapDf$cluster <- as.factor(ab$cluster)

g <- ggplot2::ggplot(umapDf,
                     ggplot2::aes(dim1,
                                  dim2,
                                  color=cluster,
                                  alpha=(DateNew)^3/(max(DateNew)^3),
                                  size=Distance))

g <- g+ggplot2::geom_point()
g
```

gganimate!
```{r}
# g <- g+ gganimate::transition_time(Date)
# g <- g+ ggplot2::labs(title="Date: {frame_time}")
# g <- g+ gganimate::shadow_mark(alpha = 0.5, size = 1)
# g <- g+ gganimate::ease_aes('linear',interval=0.001)
# g
# f<-gganimate::animate(g,duration=120)
```


```{r}
# dM <- dplyr::select(umapDf,
#                     DateNew,
#                     Vert,
#                     Distance,
#                     Cadence,
#                     Elevation,
#                     Effort,
#                     Track,
#                     Temp,
#                     HR,
#                     restDays,
#                     MaxHR)
# dM$Distance <- log10(dM$Distance)
# dM$Elevation <- log10(dM$Elevation)
# dM$Effort <- log10(dM$Effort)
# dM1 <- model.matrix(~ .^2,dM)
# dM1 <- data.matrix(dM1)
# dM1<-dM1[,apply(dM1,2,sd)>0]
# #dM1 <- scale(dM1)
# #dM1 <- scale(dM1)
# y <- foo$Pace
# 
# #res4<-spike::vbsrBootstrap(y=scale(y),x=scale(dM1),nsamp = 1000)
# glm_res<-glmnet::cv.glmnet(x=dM1,y=(y),nfolds = nrow(foo),lambda=exp(seq(-8,1,length.out=100)))
# beta1 <- glm_res$glmnet.fit$beta[,glm_res$lambda==glm_res$lambda.1se]
# beta1[beta1!=0]
# glm_res$cvm[glm_res$lambda==glm_res$lambda.1se]
# plot(glm_res)

```



Build a couple pace prediction models (model1 -> interaction, no effort, model2 -> interaction with effort, model3 -> no interaction, no effort, model4 -> no interaction, with effort)
```{r}
pace <- list()
pace$interactionNoEffort<-(lm(Pace ~ DateNew +
           (HR) +
           Vert+
           log10(Distance)+
           (Cadence)+
           log10(Elevation) +
           Vert*log10(Distance) +
           HR*Cadence+
           Cadence*Vert +
           Track + 
           restDays,foo))

pace$interactionEffort <-(lm(Pace ~ DateNew +
           (HR) +
           Vert+
           log10(Distance)+
           (Cadence)+
           log10(Elevation) +
           Vert*log10(Distance) +
           HR*Cadence+
           Cadence*Vert +
           Track + 
           restDays +
           log10(Effort),foo))

pace$noInteractionNoEffort <-(lm(Pace ~ DateNew +
           (HR) +
           Vert+
           log10(Distance)+
           (Cadence)+
           log10(Elevation) +
           Track + 
           restDays,foo))

pace$noInteractionEffort <-(lm(Pace ~ DateNew +
            (HR) +
            Vert +
            log10(Distance)+
            (Cadence)+
            log10(Elevation) +
            Track + 
            restDays + 
            log10(Effort),foo))
```

Marathon prediction
```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(882,4)),
       HR=as.list(rep(171,4)),
       Distance=as.list(rep(26.2,4)),
       Cadence=as.list(rep(180,4)),
       Vert=as.list(rep(1,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,350,NULL,350))*(26.2/60)
```

1/2 Marathon prediction
```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(882,4)),
       HR=as.list(rep(178,4)),
       Distance=as.list(rep(13.1,4)),
       Cadence=as.list(rep(182,4)),
       Vert=as.list(rep(1,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,200,NULL,200))*(13.1)
```


5k prediction
```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(882,4)),
       HR=as.list(rep(185,4)),
       Distance=as.list(rep(3.1,4)),
       Cadence=as.list(rep(186,4)),
       Vert=as.list(rep(1,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,67,NULL,67))*(3.1)
```


10k prediction
```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(882,4)),
       HR=as.list(rep(184,4)),
       Distance=as.list(rep(6.2,4)),
       Cadence=as.list(rep(184,4)),
       Vert=as.list(rep(1,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,130,NULL,130))*(6.2)
```

Thunder cougar falcon bird classic prediction

```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(396+365,4)),
       HR=as.list(rep(165,4)),
       Distance=as.list(rep(14.33,4)),
       Cadence=as.list(rep(170,4)),
       Vert=as.list(rep(4785,4)),
       Elevation=as.list(rep(400,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,200,NULL,200))*(14.33/60)
```
Rattlesnake two-fer prediction

```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(882,4)),
       HR=as.list(rep(165,4)),
       Distance=as.list(rep(19.35,4)),
       Cadence=as.list(rep(171,4)),
       Vert=as.list(rep(5489,4)),
       Elevation=as.list(rep(1000,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,362,NULL,362))*(19.35/60)
```

Yakima 25k

```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(882,4)),
       HR=as.list(rep(165,4)),
       Distance=as.list(rep(15.36,4)),
       Cadence=as.list(rep(161,4)),
       Vert=as.list(rep(5157,4)),
       Elevation=as.list(rep(1000,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,200,NULL,200))*(15.36/60)
```

squak-octogon
```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(882,4)),
       HR=as.list(rep(147,4)),
       Distance=as.list(rep(28.11,4)),
       Cadence=as.list(rep(172,4)),
       Vert=as.list(rep(9501,4)),
       Elevation=as.list(rep(500,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,390,NULL,390))*(28.11/60)
```

Orcas
```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(882,4)),
       HR=as.list(rep(183,4)),
       Distance=as.list(rep(15.5,4)),
       Cadence=as.list(rep(168,4)),
       Vert=as.list(rep(4500,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,596,NULL,596))*(15.5/60)
```


```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(882,4)),
       HR=as.list(rep(155,4)),
       Distance=as.list(rep(50,4)),
       Cadence=as.list(rep(172,4)),
       Vert=as.list(rep(8200,4)),
       Elevation=as.list(rep(1700,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,400,NULL,400))*(50/60)
```
```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(882,4)),
       HR=as.list(rep(154,4)),
       Distance=as.list(rep(62.1,4)),
       Cadence=as.list(rep(174,4)),
       Vert=as.list(rep(5190,4)),
       Elevation=as.list(rep(3000,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,500,NULL,500))*(62.1/60)
```

```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(882,4)),
       HR=as.list(rep(195,4)),
       Distance=as.list(rep(1,4)),
       Cadence=as.list(rep(186,4)),
       Vert=as.list(rep(1,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,30,NULL,30))
```