---
title: "Ultra training notebook"
output: html_notebook
---

Load data and convert date

```{r}
foo <- data.table::fread('data/ultra_training.csv')
foo$Date <- as.Date(foo$Date,tryFormats=c("%m/%d/%y"))
foo <- foo[!is.na(foo$AirTemp),]
```

Create a numeric Date variable starting at the commencement of serious training (April 9, 2019), as well as a rest days variabe.

```{r}
foo$DateNew <- as.numeric(foo$Date)-min(as.numeric(foo$Date))
foo$restDays <- c(7,diff(foo$DateNew))
foo$restDays2 <- foo$restDays
foo$restDays2[foo$restDays>6] <- 7
foo$restDays2 <- as.factor(foo$restDays2)
```


impute missing data (HR, Effort,MaxHR,EPOC)
```{r}
hr_mod <- lm(log10(HR)~log10(Distance)+Pace+GAP+Vert+Cadence+Elevation+Track+Temp+StartTime+ElapsedTime+DateNew+restDays,foo)
effort_mod <- lm(log10(Effort)~log10(Distance)+Pace+GAP+Vert+Cadence+Elevation+Track+Temp+StartTime+ElapsedTime+DateNew+restDays,foo)
MaxHR_mod <- lm(log10(MaxHR)~log10(Distance)+Pace+GAP+Vert+Cadence+Elevation+Track+Temp+StartTime+ElapsedTime+DateNew+restDays,foo)
epoc_mod <- lm(log10(EPOC)~log10(Distance)+Pace+GAP+Vert+Cadence+Elevation+Track+Temp+StartTime+ElapsedTime+DateNew+restDays,foo)
hr_pred <- predict(hr_mod,foo)
effort_pred <- predict(effort_mod,foo)
MaxHR_pred <- predict(MaxHR_mod,foo)
epoc_pred <- predict(epoc_mod,foo)
foo$HR[is.na(foo$HR)] <- 10^hr_pred[is.na(foo$HR)]
foo$Effort[is.na(foo$Effort)] <- 10^effort_pred[is.na(foo$Effort)]
foo$MaxHR[is.na(foo$MaxHR)] <- 10^MaxHR_pred[is.na(foo$MaxHR)]
foo$EPOC[is.na(foo$EPOC)] <- 10^epoc_pred[is.na(foo$EPOC)]
```


Create beats per mile and beats per grade adjusted mile stats
```{r}
foo$bpgam <- foo$GAP*foo$HR
foo$bpmile <- foo$Pace*foo$HR
```

Fit model without effort or DateNew to see progression.

```{r}
model1 <- lme4::lmer(log10(bpmile) ~ Vert +
                       log10(Distance) + 
                       Cadence +
                       Vert*log10(Distance) + 
                       Cadence*Vert +
                       Track +
                       MaxHR +
                       Humidity +
                       AirTemp +
                       Humidity*AirTemp +
                       (1|WindDirection)+
                       (1|weather),foo)

foo$resid <- scale(residuals(model1))
g <- ggplot2::ggplot(foo,ggplot2::aes(x=Date,y=resid))

g <- g + ggplot2::geom_smooth(method=MASS::rlm)

g <- g + ggplot2::geom_point()

g
```

Look at beats per grade adjusted mile

```{r}
summary(MASS::rlm((bpgam) ~ DateNew,foo))
g <- ggplot2::ggplot(foo,ggplot2::aes(x=Date,
                                      y=log10(bpgam)))

g <- g + ggplot2::geom_smooth(method=MASS::rlm)

g <- g + ggplot2::geom_point()

g
```

Fit moving average to residuals
```{r}
ma <- smooth::sma(residuals(model1),order=200)
foo$fitted <- ma$fitted
dd <- .16
g <- ggplot2::ggplot(foo,ggplot2::aes(x=Date,y=0.80/exp(exp(fitted/dd)-exp(max(fitted)/dd))))
g <- g+ ggplot2::geom_line()
g <- g+ ggplot2::scale_x_date(date_breaks='45 weeks')
g <- g+ ggplot2::ylab('Ultra Signup Ranking')
g <- g+ggplot2::geom_hline(yintercept = 0.85)

g <- g + cowplot::theme_cowplot()
g

g <- ggplot2::ggplot(foo,ggplot2::aes(x=Date,y=fitted))
g <- g+ ggplot2::geom_line()
g <- g+ ggplot2::scale_x_date(date_breaks='30 weeks')
g <- g + ggplot2::ylab('fitness')
g

cat('fitness I need to get to for 80%: ',dd*log(log(0.8)-log(0.8)+exp(max(foo$fitted)/dd)),'\n')
cat('fitness I need to get to for 82.5%: ',dd*log(log(0.8)-log(0.825)+exp(max(foo$fitted)/dd)),'\n')
cat('fitness I need to get to for 85%: ',dd*log(log(0.8)-log(0.85)+exp(max(foo$fitted)/dd)),'\n')
cat('fitness I need to get to for 87.5%: ',dd*log(log(0.8)-log(0.875)+exp(max(foo$fitted)/dd)),'\n')
cat('fitness I need to get to for 90%: ',dd*log(log(0.8)-log(0.9)+exp(max(foo$fitted)/dd)),'\n')
cat('fitness I need to get to for 92.5%: ',dd*log(log(0.8)-log(0.925)+exp(max(foo$fitted)/dd)),'\n')
cat('fitness I need to get to for 95%: ',dd*log(log(0.8)-log(0.95)+exp(max(foo$fitted)/dd)),'\n')
cat('fitness I need to get to for 97.5%: ',dd*log(log(0.8)-log(0.975)+exp(max(foo$fitted)/dd)),'\n')
cat('fitness I need to get to for 99%: ',dd*log(log(0.8)-log(0.99)+exp(max(foo$fitted)/dd)),'\n')
cat('E[USR]: ',0.80/exp(exp(foo$fitted[nrow(foo)]/dd)-exp(max(foo$fitted)/dd)),'\n')
```

```{r}
# model1list <- vector('list',nrow(foo)-30)
# malist <- vector('list',nrow(foo)-30)
# for (i in 31:nrow(foo)){
#   model1list[[i-30]] <- lm(bpmile ~ Vert +
#                   log10(Distance) +
#                   Cadence+log10(Elevation) + 
#                   Vert*log10(Distance) + 
#                   Cadence*Vert + 
#                   Track + 
#                   restDays +
#                   MaxHR + 
#                   Poles,foo[1:i])
#    malist[[i-30]]<-smooth::sma((model1list[[i-30]]$resid),ic = 'BICc')
#    print(i)
# }
```



Fit model without effort to see effects of each variable.
```{r}
model2 <- lm(log10(bpmile) ~ Vert + 
               log10(Distance) + 
               Cadence + 
               log10(Elevation) + 
               Vert*log10(Distance) + 
               Cadence*Vert + 
               Track + 
               DateNew +
               restDays + 
               MaxHR + 
               Poles,foo)
summary(model2)
foo$residFull <- scale(model2$residuals)
```


Fit model with effort to see full effects
```{r}
model3 <- lm(log10(bpmile) ~ Vert + 
               log10(Distance) + 
               Cadence +
               log10(Elevation) + 
               Vert*log10(Distance) + 
               Cadence*Vert + 
               Track + 
               log10(Effort) + 
               DateNew +
               restDays + 
               MaxHR + 
               Poles,foo)
summary(model3)
foo$residEffort <- scale(model3$residuals)
foo$pEffort <- pchisq(foo$residEffort^2,1,lower.tail=F)
```
```{r}
model3humid <- lm(log10(bpmile) ~ Vert + 
               log10(Distance) + 
               Cadence +
               log10(Elevation) + 
               Vert*log10(Distance) + 
               Cadence*Vert + 
               Track + 
               log10(Effort) + 
               DateNew +
               restDays + 
               MaxHR +
               Poles +
               StartTime
               ,foo)
summary(model3humid)
```

```{r}
model4 <- lm(log10(bpmile) ~ Vert + 
               log10(Distance) + 
               Cadence +
               log10(Elevation) + 
               Vert*log10(Distance) + 
               Cadence*Vert + 
               Track + 
               StartTime +
               restDays + 
               MaxHR + 
               Poles,foo)
summary(model4)
```


Look at the full matrix

```{r}
dplyr::arrange(foo,dplyr::desc(Date))
```

```{r}
fxn2 <- function(day,df){
  searchDays <- day - c(1:21)
  ab <- dplyr::filter(df,Date%in%searchDays)
  if(nrow(ab)>0){
    return(sum(ab$ElapsedTime))
  }else{
    return(0)
  }
}

fxn3 <- function(day,df){
  searchDays <- day - c(1:21)
  ab <- dplyr::filter(df,Date%in%searchDays)
  if(nrow(ab)>0){
    return(sum(ab$Distance))
  }else{
    return(0)
  }
}

fxn4 <- function(day,df){
  searchDays <- day - c(1:21)
  ab <- dplyr::filter(df,Date%in%searchDays)
  if(nrow(ab)>0){
    return(sum(ab$Effort))
  }else{
    return(0)
  }
}

foo$cumulativeTime <- sapply(foo$Date,fxn2,foo)
foo$cumulativeDistance <- sapply(foo$Date,fxn3,foo)
foo$cumulativeEffort <- sapply(foo$Date,fxn4,foo)
modelx <- lm(log10(bpmile) ~ Vert + 
               log10(Distance) + 
               Cadence +
               log10(Elevation) + 
               Vert*log10(Distance) + 
               log10(Effort) +
               Cadence*Vert + 
               Track + 
               DateNew +
               restDays + 
               MaxHR + 
               log10(cumulativeDistance+1) +
               log10(cumulativeTime+1)+
               log10(cumulativeEffort+1),foo)
summary(modelx)
```

plot run types

```{r}
set.seed(1)
dM <- dplyr::select(foo,
                    DateNew,
                    Vert,
                    Distance,
                    Cadence,
                    Elevation,
                    Effort,
                    Track,
                    Temp,
                    HR,
                    MaxHR,
                    StartTime,
                    ElapsedTime,
                    AirTemp,
                    Humidity,
                    WindSpeed,
                    restDays,
                    Poles,
                    EPOC)

dM$Distance <- log10(dM$Distance)
dM$Elevation <- log10(dM$Elevation)
dM$Effort <- log10(dM$Effort)
dM$AirTemp <- log10(dM$AirTemp)
dM$Humidity <- log10(dM$Humidity)
dM$WindSpeed <- log10(dM$WindSpeed)
dM <- dM[!is.na(dM$AirTemp),]
res2 <- umap::umap(scale(dM))

umapDf <- data.frame(dim1=res2$layout[,1],
                     dim2=res2$layout[,2],
                     stringsAsFactors = F)

umapDf <- cbind(umapDf,foo[!is.na(foo$AirTemp)])

ab<-kmeans(umapDf[,c('dim1','dim2')],4,nstart=1000)
umapDf$cluster <- as.factor(ab$cluster)

g <- ggplot2::ggplot(umapDf,
                     ggplot2::aes(dim1,
                                  dim2,
                                  color=cluster,
                                  alpha=(DateNew)^3/(max(DateNew)^3),
                                  size=Distance))

g <- g+ggplot2::geom_point()
g
```


Define a pace prediction function
```{r}
pacePredictor <- function(modFit,
                          predDf){
  res <- predict(modFit,
                 predDf,
                 interval='prediction',
                 level = .68)
  res2 <- predict(modFit,
                 predDf,
                 interval='prediction',
                 level = .9)
  #names(res) <- c('Prediction','Lower IQR','Upper IQR')
  #names(res2) <- c('Prediction','Lower Decile','Upper Decile')
  
  res4 <- cbind(res2[,2],res[,2],res[,1],res[,3],res2[,3])
  colnames(res4) <- c('Lower Decile','Lower IQR','Prediction','Upper IQR','Upper Decile')
  return(10^res4)
}
```


Define pace model:

```{r}
pace2 <-(lm(log10(Pace) ~ DateNew +
           (HR) +
           Vert+
           log10(Distance)+
           (Cadence)+
           Vert*log10(Distance)+
           Cadence*Vert +
           Track + 
           log10(Effort)+
           MaxHR +
           Poles + 
           MaxHR*Vert,foo))
summary(pace2)
```


Define times:

```{r}
today <- foo$DateNew[nrow(foo)]
benchmark_data <- data.table::fread('data/benchmarks.csv',data.table=F)
benchmark_data$DateNew <- today
bench2 <- dplyr::select(benchmark_data,
                        DateNew,
                        HR,
                        Vert,
                        Distance,
                        Cadence,
                        Elevation,
                        Track,
                        MaxHR,
                        Humidity,
                        AirTemp,
                        Effort,
                        Poles,
                        AirTemp,
                        Humidity)

convertTimes <- function(x){
  return(as.character(round(lubridate::seconds_to_period(x))))
}

predPaces<-pacePredictor(pace2,bench2)
predPaces<-predPaces*benchmark_data$Distance*60
predPaces[benchmark_data$Ascent,]<-predPaces[benchmark_data$Ascent,]*(1.1)
predPaces<-apply(predPaces,2,convertTimes)

benchmark_data <- cbind(benchmark_data[,1],predPaces,benchmark_data)
benchmark_data

```


impute the Power!

```{r}
power_mod <- lm(Power ~ HR + Cadence + Vert + log10(Distance),foo)
summary(power_mod)
pred1 <- predict(power_mod,foo)
foo2 <- foo
foo2$predPower <- pred1
foo2$PowerPerStep <- foo2$predPower/foo2$Cadence
foo2$PowerBerBeat <- log10(foo2$predPower) - log10(foo2$HR)
```

```{r}
set.seed(1)
ns <- 2
mod <- depmixS4::depmix(response = resid~1,
                        data=foo,
                        nstates=ns,
                        trstart = runif(ns^2))
fm <- depmixS4::fit(mod, 
          emc=depmixS4::em.control(rand=FALSE))
fm
#for(i in 1:ns){
#  print(i)
#  print(c(fm@response[[i]][[1]]@parameters))
  #plot(fm@posterior[[i]])
#}
```
1672







