---
title: "Ultra sandbox code chunks"
output: html_notebook
---

load data first


Load data and convert date

```{r}
foo <- data.table::fread('data/ultra_training.csv',data.table=F)
foo$Date <- as.Date(foo$Date,tryFormats=c("%m/%d/%y"))
```

Create a numeric Date variable starting at the commencement of serious training (April 9, 2019), as well as a rest days variabe.

```{r}
foo$DateNew <- as.numeric(foo$Date)-min(as.numeric(foo$Date))
foo$restDays <- c(8,diff(foo$DateNew))
foo$restDays2 <- log10(foo$restDays+1)
```


impute missing heart rate
```{r}
hr_mod <- lm(log10(HR)~log10(Distance)+
               Pace+
               GAP+
               Vert+
               Cadence+
               Elevation+
               Track+
               Temp+
               StartTime+
               ElapsedTime+
               DateNew+
               restDays2,foo)
hr_pred <- predict(hr_mod,foo)
foo$HR[is.na(foo$HR)] <- 10^hr_pred[is.na(foo$HR)]
summary(hr_mod)
```

impute missing effort
```{r}
effort_mod <- lm(log10(Effort)~log10(Distance)+
                   Pace+
                   GAP+
                   Vert+
                   Cadence+
                   Elevation+
                   Track+
                   Temp+
                   StartTime+
                   ElapsedTime+
                   DateNew+
                   restDays2,foo)
effort_pred <- predict(effort_mod,foo)
foo$Effort[is.na(foo$Effort)] <- 10^effort_pred[is.na(foo$Effort)]
summary(effort_mod)
```

impute maximum heart rate
```{r}
MaxHR_mod <- lm(log10(MaxHR)~log10(Distance)+
                  Pace+
                  GAP+
                  Vert+
                  Cadence+
                  Elevation+
                  Track+
                  Temp+
                  StartTime+
                  ElapsedTime+
                  DateNew+
                  restDays2,foo)
MaxHR_pred <- predict(MaxHR_mod,foo)
foo$MaxHR[is.na(foo$MaxHR)] <- 10^MaxHR_pred[is.na(foo$MaxHR)]
summary(MaxHR_mod)
```

impute epoc
```{r}
epoc_mod <- lm(log10(EPOC)~log10(Distance)+
                 Pace+
                 GAP+
                 Vert+
                 Cadence+
                 Elevation+
                 Track+
                 Temp+
                 StartTime+
                 ElapsedTime+
                 DateNew+
                 restDays2,foo)
epoc_pred <- predict(epoc_mod,foo)
foo$EPOC[is.na(foo$EPOC)] <- 10^epoc_pred[is.na(foo$EPOC)]
summary(epoc_mod)
```

Create beats per mile and beats per grade adjusted mile stats
```{r}
foo$bpgam <- foo$GAP*foo$HR
foo$bpmile <- foo$Pace*foo$HR
```


impute power
```{r}
power_mod <- lm(log10(Power)~log10(Distance)+
                 Pace+
                 GAP+
                 Vert+
                 Cadence+
                 Elevation+
                 Track+
                 Temp+
                 StartTime+
                 ElapsedTime+
                 HR+
                 restDays2,foo)
power_pred <- predict(power_mod,foo)
foo$Power[is.na(foo$Power)] <- 10^power_pred[is.na(foo$Power)]
summary(power_mod)
```

impute AirPower
```{r}
airpower_mod <- lm(log10(AirPower+1)~EPOC+
                 Temp+
                 Vert+
                 StartTime+
                 Cadence,foo)
airpower_pred <- predict(airpower_mod,foo)
foo$AirPower[is.na(foo$AirPower)] <- 10^airpower_pred[is.na(foo$AirPower)]-1
summary(airpower_mod)
```

impute FormPower
```{r}
formpower_mod <- lm(log10(FormPower)~log10(Distance)+
                 Pace+
                 GAP+
                 Cadence+
                 ElapsedTime+
                 HR,foo)
formpower_pred <- predict(formpower_mod,foo)
foo$FormPower[is.na(foo$FormPower)] <- 10^formpower_pred[is.na(foo$FormPower)]
summary(formpower_mod)
```


impute GroundTime
```{r}
groundtime_mod <- lm(log10(GroundTime)~log10(Distance)+
                 GAP+
                 Cadence+
                 Track+
                 Temp+
                 ElapsedTime+
                 HR,foo)
groundtime_pred <- predict(groundtime_mod,foo)
foo$GroundTime[is.na(foo$GroundTime)] <- 10^groundtime_pred[is.na(foo$GroundTime)]
summary(groundtime_mod)
```

impute LSS
```{r}
lss_mod <- lm(log10(LSS)~GAP+
                 Vert+
                 Cadence+
                 StartTime+
                 restDays2+
                 EPOC,foo)
lss_pred <- predict(lss_mod,foo)
foo$LSS[is.na(foo$LSS)] <- 10^lss_pred[is.na(foo$LSS)]
summary(lss_mod)
```


impute VertOsc
```{r}
vertosc_mod <- lm(log10(VertOsc)~Pace+
                 Vert+
                   GAP+
                 log10(Effort)+
                 Elevation+
                 ElapsedTime+
                 EPOC+
                 restDays2,foo)
vertosc_pred <- predict(vertosc_mod,foo)
foo$VertOsc[is.na(foo$VertOsc)] <- 10^vertosc_pred[is.na(foo$VertOsc)]
summary(vertosc_mod)
```


impute Recovery Points
```{r}
rp_mod <- lm((RecoveryPoints)~Cadence+
                 Elevation+
                 MaxHR+
                 GroundTime+
                 AirPower+
                 Pace+
                 Effort+
                 HR+
                 DateNew,foo)
rp_pred <- predict(rp_mod,foo)
foo$RecoveryPoints[is.na(foo$RecoveryPoints)] <- rp_pred[is.na(foo$RecoveryPoints)]
summary(rp_mod)
```

[1] "Date"           "Distance"       "HR"             "Pace"          
 [5] "GAP"            "Vert"           "Effort"         "Cadence"       
 [9] "Elevation"      "Track"          "Temp"           "MaxHR"         
[13] "StartTime"      "ElapsedTime"    "Power"          "FormPower"     
[17] "GroundTime"     "LSS"            "VertOsc"        "AirPower"      
[21] "AirTemp"        "Humidity"       "WindSpeed"      "WindDirection" 
[25] "weather"        "Poles"          "EPOC"           "Watch"         
[29] "RecoveryPoints" "DateNew"        "restDays"       "restDays2"     
[33] "bpgam"          "bpmile"





```{r}
whiteList<-apply(is.na(foo),2,sum)
whiteList['Date']<-1
whiteList['Watch']<-1
keepCols <- names(whiteList[whiteList==0])
predMat <- foo[,keepCols]
predMat$Track <- as.numeric(predMat$Track)
predMat$Poles <- as.numeric(predMat$Poles)
```

Power
```{r}
set.seed(1)
lambda <- seq(-8,1,length.out=100)
y<-log10(foo$Power)
wkeep <- which(!is.na(y))
y<-y[wkeep]
#y <- scale(y)
x<-data.matrix(predMat)[wkeep,]
dM1 <- model.matrix(~ .^2,predMat[wkeep,])
dM1 <- data.matrix(dM1)
wposvar <- apply(dM1,2,sd)>0
dM1<-dM1[,wposvar]
dM2 <- model.matrix(~ .^2,predMat)
dM2 <- data.matrix(dM2)
dM2 <- dM2[,wposvar]
#x2 <- model.matrix()
res1 <- glmnet::cv.glmnet(x=x,y=y,lambda = exp(lambda))
res2 <- glmnet::cv.glmnet(x=dM1,y=y,lambda=exp(lambda))
power_pred1 <- predict(res1,data.matrix(predMat))
power_pred2 <- predict(res2,dM2)
foo$Power[is.na(foo$Power)] <- 10^power_pred2[is.na(foo$Power)]
```

AirPower
```{r}
set.seed(1)
lambda <- seq(-8,-4,length.out=100)
y<-(foo$AirPower)
wkeep <- which(!is.na(y))
y<-y[wkeep]
#y <- scale(y)
x<-data.matrix(predMat)[wkeep,]
dM1 <- model.matrix(~ .^2,predMat[wkeep,])
dM1 <- data.matrix(dM1)
wposvar <- apply(dM1,2,sd)>0
dM1<-dM1[,wposvar]
dM2 <- model.matrix(~ .^2,predMat)
dM2 <- data.matrix(dM2)
dM2 <- dM2[,wposvar]
#x2 <- model.matrix()
res1 <- glmnet::cv.glmnet(x=x,y=y,lambda = exp(lambda))
res2 <- glmnet::cv.glmnet(x=dM1,y=y,lambda=exp(lambda))
power_pred1 <- predict(res1,data.matrix(predMat))
power_pred2 <- predict(res2,dM2)
foo$AirPower[is.na(foo$AirPower)] <- power_pred1[is.na(foo$AirPower)]
```

FormPower
```{r}
y<-foo$FormPower
wkeep <- which(!is.na(y))
y<-y[wkeep]
y <- scale(y)
x<-data.matrix(predMat)[wkeep,]
dM1 <- model.matrix(~ .^2,predMat[wkeep,])
dM1 <- data.matrix(dM1)
dM1<-dM1[,apply(dM1,2,sd)>0]
#x2 <- model.matrix()
res1 <- glmnet::cv.glmnet(x=x,y=y)
res2 <- glmnet::cv.glmnet(x=dM1,y=y)
```

GroundTime
```{r}
y<-foo$GroundTime
wkeep <- which(!is.na(y))
y<-y[wkeep]
y <- scale(y)
x<-data.matrix(predMat)[wkeep,]
dM1 <- model.matrix(~ .^2,predMat[wkeep,])
dM1 <- data.matrix(dM1)
dM1<-dM1[,apply(dM1,2,sd)>0]
#x2 <- model.matrix()
res1 <- glmnet::cv.glmnet(x=x,y=y)
res2 <- glmnet::cv.glmnet(x=dM1,y=y)
```

LSS
```{r}
y<-foo$LSS
wkeep <- which(!is.na(y))
y<-y[wkeep]
y <- scale(y)
x<-data.matrix(predMat)[wkeep,]
dM1 <- model.matrix(~ .^2,predMat[wkeep,])
dM1 <- data.matrix(dM1)
dM1<-dM1[,apply(dM1,2,sd)>0]
#x2 <- model.matrix()
res1 <- glmnet::cv.glmnet(x=x,y=y)
res2 <- glmnet::cv.glmnet(x=dM1,y=y)
```

VertOsc
```{r}
set.seed(1)
y<-foo$VertOsc
wkeep <- which(!is.na(y))
y<-y[wkeep]
y <- scale(y)
x<-data.matrix(predMat)[wkeep,]
dM1 <- model.matrix(~ .^2,predMat[wkeep,])
dM1 <- data.matrix(dM1)
dM1<-dM1[,apply(dM1,2,sd)>0]
#x2 <- model.matrix()
res1 <- glmnet::cv.glmnet(x=x,y=y)
res2 <- glmnet::cv.glmnet(x=dM1,y=y)
```

RecoveryPoints
```{r}
set.seed(1)
y<-foo$RecoveryPoints
wkeep <- which(!is.na(y))
y<-y[wkeep]
y <- scale(y)
x<-data.matrix(predMat)[wkeep,]
dM1 <- model.matrix(~ .^2,predMat[wkeep,])
dM1 <- data.matrix(dM1)
dM1<-dM1[,apply(dM1,2,sd)>0]
#x2 <- model.matrix()
res1 <- glmnet::cv.glmnet(x=x,y=y)
res2 <- glmnet::cv.glmnet(x=dM1,y=y)
```

AirTemp
```{r}
set.seed(1)
y<-foo$AirTemp
wkeep <- which(!is.na(y))
y<-y[wkeep]
y <- scale(y)
x<-data.matrix(predMat)[wkeep,]
dM1 <- model.matrix(~ .^2,predMat[wkeep,])
dM1 <- data.matrix(dM1)
dM1<-dM1[,apply(dM1,2,sd)>0]
#x2 <- model.matrix()
res1 <- glmnet::cv.glmnet(x=x,y=y)
res2 <- glmnet::cv.glmnet(x=dM1,y=y)
```

Humidity
```{r}
set.seed(1)
y<-foo$Humidity
wkeep <- which(!is.na(y))
y<-y[wkeep]
y <- scale(y)
x<-data.matrix(predMat)[wkeep,]
dM1 <- model.matrix(~ .^2,predMat[wkeep,])
dM1 <- data.matrix(dM1)
dM1<-dM1[,apply(dM1,2,sd)>0]
#x2 <- model.matrix()
res1 <- glmnet::cv.glmnet(x=x,y=y)
res2 <- glmnet::cv.glmnet(x=dM1,y=y)
```

WindSpeed
```{r}
set.seed(1)
y<-foo$WindSpeed
wkeep <- which(!is.na(y))
y<-y[wkeep]
y <- scale(y)
x<-data.matrix(predMat)[wkeep,]
dM1 <- model.matrix(~ .^2,predMat[wkeep,])
dM1 <- data.matrix(dM1)
dM1<-dM1[,apply(dM1,2,sd)>0]
#x2 <- model.matrix()
res1 <- glmnet::cv.glmnet(x=x,y=y)
res2 <- glmnet::cv.glmnet(x=dM1,y=y)
```


weather
```{r}
set.seed(1)
y<-as.factor(foo$weather)
wkeep <- which(!is.na(y))
y<-y[wkeep]
#y <- scale(y)
x<-data.matrix(predMat)[wkeep,]
dM1 <- model.matrix(~ .^2,predMat[wkeep,])
dM1 <- data.matrix(dM1)
dM1<-dM1[,apply(dM1,2,sd)>0]
#x2 <- model.matrix()
res1 <- glmnet::cv.glmnet(x=x,y=y,family='multinomial')
#res2 <- glmnet::cv.glmnet(x=dM1,y=y,family='multinomial')
```
WindDirection
```{r}
set.seed(1)
y<-as.factor(foo$WindDirection)
wkeep <- which(!is.na(y))
y<-y[wkeep]
#y <- scale(y)
x<-data.matrix(predMat)[wkeep,]
dM1 <- model.matrix(~ .^2,predMat[wkeep,])
dM1 <- data.matrix(dM1)
dM1<-dM1[,apply(dM1,2,sd)>0]
#x2 <- model.matrix()
res1 <- glmnet::cv.glmnet(x=x,y=y,family='multinomial')
#res2 <- glmnet::cv.glmnet(x=dM1,y=y,family='multinomial')
```


gganimate!
```{r}
# g <- g+ gganimate::transition_time(Date)
# g <- g+ ggplot2::labs(title="Date: {frame_time}")
# g <- g+ gganimate::shadow_mark(alpha = 0.5, size = 1)
# g <- g+ gganimate::ease_aes('linear',interval=0.001)
# g
# f<-gganimate::animate(g,duration=120)
```



predictive modeling noodling
```{r}
dM <- dplyr::select(foo,
                    DateNew,
                    Vert,
                    Distance,
                    Cadence,
                    Elevation,
                    Effort,
                    Track,
                    HR,
                    restDays,
                    MaxHR,
                    AirTemp,
                    Humidity,
                    WindSpeed,
                    Poles)
dM$Distance <- log10(dM$Distance)
dM$Elevation <- log10(dM$Elevation)
dM$Effort <- log10(dM$Effort)
dM$AirTemp <- log10(dM$AirTemp)
dM$Humidity <- log10(dM$Humidity)
dM$WindSpeed <- log10(dM$WindSpeed)
dM <- dM[!is.na(dM$AirTemp),]
foo2 <- foo[!is.na(foo$AirTemp),]
dM1 <- model.matrix(~ .^2,dM)
dM1 <- data.matrix(dM1)
dM1<-dM1[,apply(dM1,2,sd)>0]
#dM1 <- scale(dM1)
#dM1 <- scale(dM1)
y <- foo2$Pace
set.seed(9)
res4<-spike::vbsrBootstrap(y=scale(y),x=scale(dM1),nsamp = 100)



dM2 <- dM1%*%diag(res4[-1])
colnames(dM2) <- colnames(dM1)
dM2 <- dM2[,apply(dM2,2,sd)>0]

spike_boot<-(lm(y~dM2))
spike_boot_beta<- summary(spike_boot)$coef[,1]


# glm_res<-glmnet::cv.glmnet(x=(dM1),y=scale(y),nfolds = nrow(foo),lambda=exp(seq(-16,1,length.out=100)),alpha=0)
# beta1 <- glm_res$glmnet.fit$beta[,glm_res$lambda==glm_res$lambda.1se]
# beta1[beta1!=0]
# beta2 <- glm_res$glmnet.fit$beta[,glm_res$lambda==glm_res$lambda.min]
# beta2[beta2!=0]
# glm_res$cvm[glm_res$lambda==glm_res$lambda.1se]
# glm_res$cvm[glm_res$lambda==glm_res$lambda.min]
# plot(glm_res)

```
```{r}
mod44<-(lm(Pace ~ Cadence +
           HR +
           log10(Effort)+
           Vert+
           Vert*log10(AirTemp)+
           DateNew*Vert+
           Cadence*HR+
           Cadence*log10(Distance)+
           log10(Distance)*HR+
           DateNew*Track+
           Vert*log10(Distance)+
           Vert*log10(Effort)+
           Vert*MaxHR+
           Vert*Cadence -
           DateNew-
           log10(AirTemp)-
           Track-
           log10(Distance)-
           MaxHR,foo1))
summary(mod44)
```

```{r}
mod44<-(lm(Pace ~ Vert +
           Cadence +
           DateNew*Vert+
           Vert*log10(Elevation)-
           DateNew-
           log10(Elevation),foo1))
summary(mod44)
```


old pace prediction code: Build a couple pace prediction models (model1 -> interaction, no effort, model2 -> interaction with effort, model3 -> no interaction, no effort, model4 -> no interaction, with effort)
```{r}
pace <- list()
foo1 <- foo
pace$interactionNoEffort<-(lm(Pace ~ DateNew +
           (HR) +
           Vert+
           log10(Distance)+
           (Cadence)+
           log10(Elevation) +
           Vert*log10(Distance) +
           HR*Cadence+
           Cadence*Vert +
           Track + 
           restDays + 
           MaxHR +
           log10(Humidity)+
           log10(AirTemp)+
           log10(Humidity)*log10(AirTemp),foo1))

pace$interactionEffort <-(lm(Pace ~ DateNew +
           (HR) +
           Vert+
           log10(Distance)+
           (Cadence)+
           log10(Elevation) +
           Vert*log10(Distance) +
           HR*Cadence+
           Cadence*Vert +
           Track + 
           restDays +
           log10(Effort)+
           MaxHR +
           log10(Humidity)+
           log10(AirTemp)+
           log10(Humidity)*log10(AirTemp),foo1))

pace$noInteractionNoEffort <-(lm(Pace ~ DateNew +
           (HR) +
           Vert+
           log10(Distance)+
           (Cadence)+
           log10(Elevation) +
           Track + 
           restDays+
           MaxHR +
           log10(Humidity)+
           log10(AirTemp)+
           log10(Humidity)*log10(AirTemp),foo1))

pace$noInteractionEffort <-(lm(Pace ~ DateNew +
            (HR) +
            Vert +
            log10(Distance)+
            (Cadence)+
            log10(Elevation) +
            Track + 
            restDays + 
            log10(Effort) +
            MaxHR +
            log10(Humidity)+
            log10(AirTemp)+
            log10(Humidity)*log10(AirTemp),foo1))
```

```{r}
pace2 <- list()
foo1 <- foo
pace2$interactionNoEffort<-(lm(Pace ~ DateNew +
           (HR) +
           Vert+
           log10(Distance)+
           (Cadence)+
           log10(Elevation) +
           Vert*log10(Distance) +
           HR*Cadence+
           Cadence*Vert +
           Track + 
           restDays,foo1))

pace2$interactionEffort <-(lm(Pace ~ DateNew +
           (HR) +
           Vert+
           log10(Distance)+
           (Cadence)+
           log10(Elevation) +
           Vert*log10(Distance) +
           HR*Cadence+
           Cadence*Vert +
           Track + 
           restDays +
           log10(Effort),foo1))

pace2$noInteractionNoEffort <-(lm(Pace ~ DateNew +
           (HR) +
           Vert+
           log10(Distance)+
           (Cadence)+
           log10(Elevation) +
           Track + 
           restDays,foo1))

pace2$noInteractionEffort <-(lm(Pace ~ DateNew +
            (HR) +
            Vert +
            log10(Distance)+
            (Cadence)+
            log10(Elevation) +
            Track + 
            restDays + 
            log10(Effort),foo1))
```


Marathon prediction
```{r}
today <- foo1$DateNew[nrow(foo1)]
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(171,4)),
       Distance=as.list(rep(26.2,4)),
       Cadence=as.list(rep(180,4)),
       Vert=as.list(rep(1,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,350,NULL,350))*(26.2/60)
```

Mile prediction
```{r}
today <- foo1$DateNew[nrow(foo1)]
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(183,4)),
       Distance=as.list(rep(1,4)),
       Cadence=as.list(rep(195,4)),
       Vert=as.list(rep(1,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,20,NULL,20))
```

1/2 Marathon prediction
```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(178,4)),
       Distance=as.list(rep(13.1,4)),
       Cadence=as.list(rep(182,4)),
       Vert=as.list(rep(1,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,200,NULL,200))*(13.1)
```


5k prediction
```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(185,4)),
       Distance=as.list(rep(3.1,4)),
       Cadence=as.list(rep(186,4)),
       Vert=as.list(rep(1,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,67,NULL,67))*(3.1)
```


10k prediction
```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(184,4)),
       Distance=as.list(rep(6.2,4)),
       Cadence=as.list(rep(184,4)),
       Vert=as.list(rep(1,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,130,NULL,130))*(6.2)
```

```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(183,4)),
       Distance=as.list(rep(6.3,4)),
       Cadence=as.list(rep(185,4)),
       Vert=as.list(rep(180,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(2,4)),
       Effort=list(NULL,140,NULL,140))*(6.3)
```

Thunder cougar falcon bird classic prediction

```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(165,4)),
       Distance=as.list(rep(14.33,4)),
       Cadence=as.list(rep(170,4)),
       Vert=as.list(rep(4785,4)),
       Elevation=as.list(rep(400,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,200,NULL,200))*(14.33/60)
```
Rattlesnake two-fer prediction

```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(165,4)),
       Distance=as.list(rep(19.35,4)),
       Cadence=as.list(rep(171,4)),
       Vert=as.list(rep(5489,4)),
       Elevation=as.list(rep(1000,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,362,NULL,362))*(19.35/60)
```

Yakima 25k

```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(165,4)),
       Distance=as.list(rep(15.36,4)),
       Cadence=as.list(rep(161,4)),
       Vert=as.list(rep(5157,4)),
       Elevation=as.list(rep(1000,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,200,NULL,200))*(15.36/60)
```

squak-octogon
```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(147,4)),
       Distance=as.list(rep(28.11,4)),
       Cadence=as.list(rep(172,4)),
       Vert=as.list(rep(9501,4)),
       Elevation=as.list(rep(500,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,390,NULL,390))*(28.11/60)
```

Orcas
```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(183,4)),
       Distance=as.list(rep(15.5,4)),
       Cadence=as.list(rep(168,4)),
       Vert=as.list(rep(4500,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,596,NULL,596))*(15.5/60)
```


```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(155,4)),
       Distance=as.list(rep(50,4)),
       Cadence=as.list(rep(172,4)),
       Vert=as.list(rep(8200,4)),
       Elevation=as.list(rep(1700,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,400,NULL,400))*(50/60)
```
```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(154,4)),
       Distance=as.list(rep(62.1,4)),
       Cadence=as.list(rep(174,4)),
       Vert=as.list(rep(5190,4)),
       Elevation=as.list(rep(3000,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,500,NULL,500))*(62.1/60)
```

```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(195,4)),
       Distance=as.list(rep(1,4)),
       Cadence=as.list(rep(186,4)),
       Vert=as.list(rep(1,4)),
       Elevation=as.list(rep(100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,30,NULL,30))
```

```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(161,4)),
       Distance=as.list(rep(30.96,4)),
       Cadence=as.list(rep(178,4)),
       Vert=as.list(rep(5049,4)),
       Elevation=as.list(rep(3500,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,300,NULL,300))*(30.96/60)
```

```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today+30,4)),
       HR=as.list(rep(155,4)),
       Distance=as.list(rep(51.8,4)),
       Cadence=as.list(rep(172,4)),
       Vert=as.list(rep(1600,4)),
       Elevation=as.list(rep(300,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,350,NULL,350))*(51.8/60)
```


```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(156,4)),
       Distance=as.list(rep(6.09,4)),
       Cadence=as.list(rep(180,4)),
       Vert=as.list(rep(735,4)),
       Elevation=as.list(rep(300,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,70,NULL,70))*(6.09/60)
```

```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(178,4)),
       Distance=as.list(rep(9.53,4)),
       Cadence=as.list(rep(178,4)),
       Vert=as.list(rep(3238,4)),
       Elevation=as.list(rep(600,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,250,NULL,250))*(9.53/60)
```

```{r}
mapply(pacePredictor,
       modFit=pace,
       DateNew=as.list(rep(today,4)),
       HR=as.list(rep(179,4)),
       Distance=as.list(rep(9.61,4)),
       Cadence=as.list(rep(175,4)),
       Vert=as.list(rep(3202,4)),
       Elevation=as.list(rep(1100,4)),
       restDays=as.list(rep(1,4)),
       Effort=list(NULL,264,NULL,264))*(9.61/60)
```


compute some cumulative efforts

```{r}
fxn2 <- function(day,df){
  searchDays <- day - c(1:21)
  ab <- dplyr::filter(df,Date%in%searchDays)
  if(nrow(ab)>0){
    return(sum(ab$Distance))
  }else{
    return(0)
  }
}

foo$cumulativeDistance <- sapply(foo$Date,fxn2,foo)
modelx <- lm(bpmile ~ Vert + 
               log10(Distance) + 
               Cadence +
               log10(Elevation) + 
               Vert*log10(Distance) + 
               Cadence*Vert + 
               Track + 
               DateNew +
               restDays + 
               MaxHR + 
               log10(cumulativeDistance+1),foo)
summary(modelx)
```



```{r}
foo2 <- foo
foo2$Distance <- log10(foo2$Distance)
foo2$Effort <- log10(foo2$Effort)
foo2$Humidity <- log10(foo2$Humidity)
foo2$AirTemp <- log10(foo2$AirTemp)
foo2$Elevation <- log10(foo2$Elevation)
foo2 <- foo2[!is.na(foo2$AirTemp),]
foobar <- randomForest::randomForest(Pace ~ DateNew +
           (HR) +
           Vert+
           (Distance)+
           (Cadence)+
           Elevation+
           Track + 
           (Effort)+
           MaxHR +
           Poles + 
           (Humidity) +
           (AirTemp) + 
           restDays,
           foo2)
bench2$restDays<-1
ab<-predict(foobar,bench2)
names(ab) <- benchmark_data[,1]
ab
randomForest::varImpPlot(foobar)
x1 <- dplyr::select(foo2, DateNew,
                    HR,
                    Vert,
                    Distance,
                    Cadence,
                    Elevation,
                    Track,
                    Effort,
                    MaxHR,
                    Poles,
                    Humidity,
                    AirTemp,
                    restDays)
ab2 <- spike::vbsrBootstrap(y=foo2$Pace,x=x1)
```
